Topological Equivariant Artist Model

### 2021-03-26 slides/main.tex
new figure code + new figures + animated pdf

\documentclass[xcolor={dvipsnames}, handout]{beamer}
%\documentclass[xcolor={dvipsnames}]{beamer}
%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}
\newtheorem{prop}{Proposition} %math?

@@

\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}

@@

\begin{frame}{Visualizations are structure preserving maps}
    %add arrow between data and vizualizations
    % we're focused on building good arrows 
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/intro/viz_same.png}
    \end{figure}
    \pause 
    \begin{description}
        \item[equivariance] properties of data and visual encoding match
    \pause 
        \item[continuity] connectivity of data and visual encoding match
    \end{description}
\end{frame}

\begin{frame}{Domain specific libraries assume data structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        \begin{tabulary}{\textwidth}{|L|l|l|}\toprule
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} & &\\
             \bottomrule
        \end{tabulary}
    \end{table}
\end{frame}

@@

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begin{equation*}\huge
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation*}
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
\end{frame}

@@

\begin{frame}{Structure of components of \dfiber}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
        \begin{equation*}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation*}
        \pause
        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
        \begin{equation*}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation*}
        where $\bullet$ is associative and has an identity action. 
\end{frame}

@@

\begin{frame}{Structural \textit{keys} with associated \textit{values} \cite{munznerVisualizationAnalysisDesign2014}}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
    \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
\end{figure}
\pause
\begin{figure}[H]
    \includegraphics[width=.5\textheight]{figures/math/base.png}
    \label{fig:base_example}
\end{figure}
\end{frame}

@@

\subsection{Graphics}
\begin{frame}{Graphic Bundle $(\gtotal,\,\gbase,\,\pi ,\,\gfiber)$} % watch dr explaination
    Continuity is preserved via the many \gbasepoint\ to one \dbasepoint\ map $\vindex: \gbase\rightarrow\dbase$
    \begin{equation*} 
        \begin{tikzcd}[ampersand replacement=\&]
            F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']                                 \\
                              \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[ll, "\xi"'] \arrow[u, "\rho"', bend right]
            \end{tikzcd}
        \end{equation*}
    \pause
    \gfiber\ is a proxy for the target display, for example $(x,\, y,\, r,\, g,\, b) \in \gfiber$
    \pause
    \begin{figure}[H] %%%bring up one at a time
        \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
        \label{fig:graphic_retraction_map}
    \end{figure}
\end{frame}

@@

\begin{frame}{Monoid Equivariance: Partial Orders}
    \begin{figure} %break out steps/reveal
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}

@@

\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \end{figure}   
\end{frame}

\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/line.png}
    \end{figure}
\end{frame}

@@

\begin{frame}{Build \vmark\ over \dbase:\vmarkd}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
\begin{equation*}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation*} 
\end{frame}

@@

\begin{frame}[fragile]{How do we make things with this model?}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \end{subfigure}
    \end{figure}
    \begin{columns}
    \column{0.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Point(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \column{.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Line(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \end{columns}
\end{frame}

\subsection{Visual Transforms}
\begin{frame}[fragile]{How do we implement \vchannel?}
    \begin{minted}{python}
    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                    'y': {'name': 'v2', 'encoder': lambda x: x},
                    'facecolors': {'name':'v3', 'encoder': cmap}, 
                    's':{'name': None , 
                        'encoder': lambda _: itertools.repeat(.02)}}
    \end{minted}
    \begin{itemize}
        \item \mintinline{python}{lambda x: x} is identity \vchannel\
        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
        \item \mintinline{python}{color.Categorical} is custom \vchannel
    \end{itemize}
\end{frame}

\subsection{Artist}
\begin{frame}[fragile]{Artist} %% rewrite w/ letters in talk
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, E, V, *args, **kwargs):
                # set properties that are specific to the artist
                # stash the input E and V
                super().__init__(*args, **kwargs)

@@

\begin{frame}[fragile]{Artists: Scatter \& Line \vmarkd}
\begin{minted}{python}
    class Point(mcollections.Collection):
        def assemble(self, x, y, s, facecolors='C0' ):
            # construct geometries of the circle glyphs in visual coordinates
            # set attributes of glyphs

@@

\begin{frame}[fragile]{Same Artist, Different Data Configurations}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/linec_1.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/lined_1.png}
        \end{subfigure}
    \end{figure}
\begin{minted}{python}
LineData(FB, edge_table, vertex_table, connect=True)
LineData(FB, edge_table, vertex_table, num_samples=2, connect=False)
\end{minted}
\end{frame}

@@

@@

\begin{frame}{\vfiber\ Components}    
\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \label{tab:mpl_visual_variable_fiber}
\end{table}
\end{frame}

>>>>

%\documentclass[xcolor={dvipsnames}, handout]{beamer}
\documentclass[xcolor={dvipsnames}]{beamer}
%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}
\newtheorem{prop}{Proposition} %math?

@@

\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}

@@

\begin{frame}{Visualizations are structure preserving maps}
    \begin{figure}
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/intro/table_only.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/intro/map.png}
            \onslide<3|handout:0>\includegraphics[width=1\linewidth]{figures/intro/time.png}
            \onslide<4->\includegraphics[width=1\linewidth]{figures/intro/artist.png}
        \end{overprint}
    \end{figure}
    \pause 
    \begin{description}
        \uncover<5->{\item[equivariance] properties of data and visual encoding match}
        \uncover<6>{\item[continuity] connectivity of data and visual encoding match}
    \end{description}
\end{frame}

\begin{frame}{Domain specific libraries assume data structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        %
        \begin{tabular}{>{\onslide<1->}l>{\onslide<2->}l>{\onslide<3->}l}
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau \cite{StoltePolaris2002}& &\\
            \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\

        \end{tabular}
    \end{table}
\end{frame}

@@

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begingroup 
    \Huge
    \begin{equation*}
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation*}
    \endgroup
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
\end{frame}

@@

\begin{frame}{Monoids are the structure of the components of \dfiber}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
        \begin{equation*}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation*}
        \pause
        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
        \begin{equation*}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation*}
        where $\bullet$ is associative and has an identity action. 
\end{frame}

@@

\begin{frame}{\dbase\ is the space of \textit{keys} into data in \dtotal \cite{munznerVisualizationAnalysisDesign2014}}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
    \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
\end{figure}
\begin{figure}[H]
    \includegraphics[width=.5\textheight]{figures/math/base.png}
    \label{fig:base_example}
\end{figure}
\end{frame}

@@

\subsection{Graphics}
\begin{frame}{Graphic Bundle $(\gtotal,\,\gbase,\,\pi ,\,\gfiber)$} % watch dr explaination
    \only<2->{Continuity is preserved via the many \gbasepoint\ to one \dbasepoint\ map $\vindex: \gbase\rightarrow\dbase$}

    \only<1>{
        \begin{equation*} 
            \begin{tikzcd}[ampersand replacement=\&]
                F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']              \\
                                  \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[u, "\rho"', bend right]
                \end{tikzcd}
            \end{equation*}}
    \only<2->{
    \begin{equation*} 
        \begin{tikzcd}[ampersand replacement=\&]
            F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']                                 \\
                              \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[ll, "\xi"'] \arrow[u, "\rho"', bend right]
            \end{tikzcd}
        \end{equation*}}
    \only<3->{   
    \gfiber\ is a proxy for the target display, for example $(x,\, y,\, r,\, g,\, b) \in \gfiber$}
    \pause
    \begin{figure}[H] %%%bring up one at a time
        \begin{overprint}
            \onslide<4|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_s.png}
            \onslide<5|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_s.png}
            \onslide<6>\includegraphics[width=1\linewidth]{figures/math/image_s.png}
        \end{overprint}
    \end{figure}
\end{frame}

@@

\begin{frame}{Monoid Equivariance: Partial Orders}
    \begin{figure} 
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/partial.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/partial_invalid.png}
            \onslide<3>\includegraphics[width=1\linewidth]{figures/math/partial_fixed.png}
        \end{overprint}
    \end{figure}
\end{frame}

@@

\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
    \begin{figure}[H]
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_q.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_with_s.png}
        \end{overprint}
    \end{figure}   
\end{frame}

\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
    \begin{figure}[H]
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_q.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_with_s.png}
        \end{overprint}
    \end{figure}   
\end{frame}

@@

\begin{frame}{Build \vmark\ over \dbase: \vmarkd}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
\begin{equation*}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation*} 
\end{frame}

@@

\begin{frame}[fragile]{How do we make things?}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \end{subfigure}
    \end{figure}
    \begin{columns}
    \column{0.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Point(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \column{.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Line(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \end{columns}
\end{frame}

\subsection{Visual Transforms}
\begin{frame}[fragile]{ \vchannel}
    \begin{minted}{python}
    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                    'y': {'name': 'v2', 'encoder': lambda x: x},
                    'facecolors': {'name':'v3', 'encoder': cmap}, 
                    's':{'name': None , 
                        'encoder': lambda _: itertools.repeat(.02)}}
    \end{minted}
    \begin{itemize}
        \item \mintinline{python}{lambda x: x} is identity \vchannel\
        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
        \item \mintinline{python}{color.Categorical} is custom \vchannel
    \end{itemize}
\end{frame}

\subsection{Artist}
\begin{frame}[fragile]{\vartist} %% rewrite w/ letters in talk
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, E, V, *args, **kwargs):
                # set properties that are specific to the artist
                # stash the input E and V
                super().__init__(*args, **kwargs)

@@

\begin{frame}[fragile]{\vmarkd}
\begin{minted}{python}
    class Point(mcollections.Collection):
        def assemble(self, x, y, s, facecolors='C0' ):
            # construct geometries of the circle glyphs in visual coordinates
            # set attributes of glyphs

@@

\begin{frame}[fragile]{Same Artist, Different \dtotal}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/linec_1.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/lined_1.png}
        \end{subfigure}
    \end{figure}
\begin{minted}{python}
LineData(FB, edge_table, vertex_table, connect=True)
LineData(FB, edge_table, vertex_table, num_samples=2, connect=False)
\end{minted}
\end{frame}

@@

\begin{frame}{Acknowledgments}
    \begin{itemize}
        \item Professor Michael Grossberg and Dr. Thomas Caswell
        \item Professor Haralick, Professor Vo, Professor Manovich, Dr. Hanwell
        \item Matplotlib development team 
        \item CZI EOSS (grant number 2019-207333) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
    \end{itemize}

@@

\begin{frame}{\vfiber\ Components}    
\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \label{tab:mpl_visual_variable_fiber}
\end{table}
\end{frame}
\begin{frame}{Monoid Equivariance: Partial Orders}
    \begin{figure} %break out steps/reveal
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}
